---
title: "Active courtship analysis"
output: 
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: no
    toc: no
editor_options: 
  chunk_output_type: console
header-includes: |
  \usepackage{lipsum} \usepackage{float} \floatplacement{figure}{H}
---

```{r knitsetup, include=FALSE}
knitr::opts_knit$set(root.dir='../',fig_path="../figs/")
```


```{r chunksetup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.extra='',fig.pos="H",
                      fig.path = "../figs/",
                      dpi = 300,fig.keep='none',dev='png')
```

```{r loadLibs, message=FALSE}
library(stringr)
library(dplyr)
library(scales)
library(tidyverse)
require(lme4)
library(knitr)
library(emmeans)
library(MuMIn)
library(vioplot)
library(kableExtra)
```

```{r}
sex_cols<-c(Female="#7fc97f",Male="#beaed4")
trial_pch<-c(0:7,9:10)
```

```{r}
behav_data<-read.csv("processed_data/combined_behavioural_data.csv")
```

We can summarize the overall behaviours and work out some summary statistics:

```{r}
behav_tab<-table(behav_data$behavior)
# recode "Second female" to just be "female"
behav_data$subject[behav_data$subject == "Second female"] <- "Female"
```

Of the `r nrow(behav_data)` observed behaviours, `r sum(behav_tab[c("Inactive", "Out of sight")])` are either "Inactive" or "Out of sight". Disregarding those, and our 'active' courtship behaviours make up `r sum(behav_tab[c("Pose", "Wiggle")])/(nrow(behav_data) - sum(behav_tab[c("Inactive", "Out of sight")]))*100`. Chasing and surrounding make up the next largest chunk of behaviours, taking `r sum(behav_tab[c("Chasing", "Surrounding")])/(nrow(behav_data) - sum(behav_tab[c("Inactive", "Out of sight")]))*100`.

```{r}
active_data <- behav_data[behav_data$behavior %in% c("Pose","Wiggle"),]
active_data$logDuration <- log(active_data$Duration)
active_data$logTotalCourtship <- log(active_data$total_time_of_courtship)
```

Before moving forward with the analysis, we want to restrict the analysis to behaviours that are intersexual rather than any female-female interactions.

```{r}
active_data <- active_data[which(active_data$subject != active_data$modifier_2),]
active_data <- active_data[which(active_data$modifier_2 != "None"),]
```


And we need to ensure that the factors are factors

```{r}
active_data$subject<- factor(active_data$subject, levels=c("Female","Male"))
active_data$Trial <- factor(active_data$Trial, levels=c(
  "Trial 1",
  "Trial 2",
  "Trial 3",
  "Trial 4",
  "Trial 5",
  "Trial 6",
  "Trial 7",
  "Trial 8",
  "Trial 9",
  "Trial 10"
))

active_data$Time_of_Day <- factor(active_data$Time_of_Day, levels=c(
  "AM","Noon"
))
```



## Exploratory analysis


How many hours of courtship were recorded? We'll consider the total duration of each courtship bout.

```{r}
all_courtship_durs<-tapply(behav_data$total_time_of_courtship,
                           behav_data$bout_number,
                           max)
sum(all_courtship_durs)/60/60 #convert to hrs
summary(all_courtship_durs/60) #in mins
```

```{r}
hist(all_courtship_durs/60,
     xlab="Courtship bout duration (min)",
     ylab="Number of bouts",
     main="")
```


What was the total time of active courtship behaviours?

```{r}
behav_times<-tapply(behav_data$Duration,
                    behav_data$behavior,
                    sum)
court_behavs<-c("Pose","Wiggle","Chasing","Surrounding")
sum(behav_times[court_behavs])/60/60
```

How about group sizes?

```{r}
group_sizes<-tapply(behav_data$Group_size,
                    behav_data$bout_number,
                    mean)
hist(group_sizes)
```

The average group size was `r mean(group_sizes)`, with a minimum of `r min(group_sizes)` and a maximum of `r max(group_sizes)`.

## Initiation


Which sex initiated displays more frequently?

```{r}
initiation<-as.data.frame(do.call(rbind,by(behav_data,
               behav_data$bout_number,
               function(bout_dat){ 
                 
                 active<-bout_dat[bout_dat$behavior != "Inactive",]
                 initiator<-active[1,"subject"]
                 if(length(table(active$subject))>1){
                   reciprocated<-TRUE
                 } else{
                   reciprocated<-FALSE
                 }
                 return(c(initiatior = initiator, reciprocated=reciprocated, bout=bout_dat$bout_number[1]))
               }
               )))
reciprocated<-initiation[initiation$reciprocated==TRUE,]
nrow(reciprocated)/nrow(initiation)
```
```{r}
recipInit<-table(reciprocated$initiatior)
recipInit/sum(recipInit)*100
# two-sided proportion test
res<-prop.test(x=recipInit[1], n=sum(recipInit), p = NULL, alternative = "two.sided",
               correct = TRUE)
res
```

## To whom do females display?

```{r}
# active behavs from females
Comp<- behav_data[!(behav_data$subject=="Male") & behav_data$behavior %in% c("Wiggle","Pose"),] 
# Number of female displays towards either males or females and those displays split into type of display
a <-sum(Comp$behavior == "Wiggle" & Comp$modifier_2 == "Male") 
b <-sum(Comp$behavior == "Wiggle" & Comp$modifier_2 == "Female")
c <-sum(Comp$behavior == "Pose" & Comp$modifier_2 == "Male")
d <-sum(Comp$behavior == "Pose" & Comp$modifier_2 == "Female")
total <- sum(c(a,b,c,d))
Comp <- matrix(c((a/total),(b/total),(c/total),(d/total)),ncol=2,byrow=TRUE)
colnames(Comp)<-c("Male", "Female")
rownames(Comp)<- c("Wiggle", "Pose")
Comp <- as.table(Comp*100)
kable(Comp)
```

Is there a difference between female displays towards female and males?

```{r}
prop.test(x=a+c, n=total, p = NULL, alternative = "two.sided",
               correct = TRUE)
```
```{r Fig1Displays, out.height='3.5in', out.width='5.75in', dpi=300, fig.keep='last'}
#png("./figs/Fig1_displays.png",height = 3.5,width=5.75,units = "in", res=1000)
# initiation
par(mfrow=c(1,2),mar=c(4,4,2,1))
barplot(recipInit, 
        xlab="Sex initiating courtship",
        ylab="Number of courtship events", 
        names.arg=c("Female", "Male"), 
        col=sex_cols[c("Female", "Male")],
        ylim=c(0,200))
legend(y=200,x=-0.4,as.expression(bquote(bold("A"))),cex=2,bty='n',xpd=TRUE)
barplot(Comp[,c(2,1)],
        ylab = "% of female displays",
        xlab = "Sex receiving female displays",
        ylim =c(0,100),
        density=c(0,15,0,15),
        legend=rownames(Comp),
        args.legend=list(
          x="left",
          bty='n'
        ),
        col=sex_cols["Female"]
)
legend(y=100,x=-0.4,as.expression(bquote(bold("B"))),cex=2,bty='n',xpd=TRUE)
#dev.off()
```

## Active behaviours


```{r}
active_data <- behav_data[behav_data$behavior %in% c("Pose","Wiggle"),]
active_data$logDuration <- log(active_data$Duration)
active_data$logTotalCourtship <- log(active_data$total_time_of_courtship)
```

Before moving forward with the analysis, we want to restrict the analysis to behaviours that are intersexual rather than any female-female interactions.

```{r}
active_data <- active_data[which(active_data$subject != active_data$modifier_2),]
active_data <- active_data[which(active_data$modifier_2 != "None"),]
```
And we need to ensure that the factors are factors
```{r}
active_data$subject<- factor(active_data$subject, levels=c("Female","Male"))
active_data$Trial <- factor(active_data$Trial, levels=c(
  "Trial 1",
  "Trial 2",
  "Trial 3",
  "Trial 4",
  "Trial 5",
  "Trial 6",
  "Trial 7",
  "Trial 8",
  "Trial 9",
  "Trial 10"
))
active_data$Time_of_Day <- factor(active_data$Time_of_Day, levels=c(
  "AM","Noon"
))
```

```{r}
table(active_data$behavior, active_data$subject)
```



```{r}
par(mfrow=c(2,2))
hist(active_data$Duration)
hist(active_data$logDuration)
hist(active_data$total_time_of_courtship)
hist(active_data$logTotalCourtship)
```




What are the things that we want to investigate? We first need to check if any of these variables are highly correlated.

- sex (aka 'subject')
- total time of courtship
- time of day
- day filmed
- group size
- Trial/bout_number (random effect)

```{r}
summary(active_data[,c("logDuration","subject","Time_of_Day","Day_filmed","total_time_of_courtship","Group_size","Trial","bout_number")])
```


```{r}
par(mfrow=c(1,5))
plot(as.factor(active_data$Trial), as.factor(active_data$subject))
plot(as.factor(active_data$Trial), active_data$Group_size)
plot(as.factor(active_data$Trial), active_data$Day_filmed)
plot(as.factor(active_data$Trial), active_data$total_time_of_courtship)
plot(as.factor(active_data$Trial), as.factor(active_data$Time_of_Day))
```

```{r}
par(mfrow=c(1,5))
plot(active_data$bout_number, as.factor(active_data$subject))
plot(active_data$bout_number, active_data$Group_size)
plot(active_data$bout_number, active_data$Day_filmed)
plot(active_data$bout_number, active_data$total_time_of_courtship)
plot(active_data$bout_number, as.factor(active_data$Time_of_Day))
```

```{r}
par(mfrow=c(1,4))
plot(active_data$Group_size, as.factor(active_data$subject))
plot(active_data$Group_size, active_data$Day_filmed)
plot(active_data$Group_size, active_data$total_time_of_courtship)
plot(active_data$Group_size, as.factor(active_data$Time_of_Day))
```

```{r}
par(mfrow=c(1,3))
plot(active_data$Day_filmed, as.factor(active_data$subject))
plot(active_data$Day_filmed, active_data$total_time_of_courtship)
plot(active_data$Day_filmed, as.factor(active_data$Time_of_Day))
```

```{r}
par(mfrow=c(1,2))
plot(active_data$total_time_of_courtship, as.factor(active_data$subject))
plot(active_data$total_time_of_courtship, as.factor(active_data$Time_of_Day))
```


```{r}
par(mfrow=c(1,1))
plot(as.factor(active_data$subject), as.factor(active_data$Time_of_Day))
```

There don't seem to be any major relationships between these predictors. 


## Model selection

For this analysis we will only consider bouts of courtship that were reciprocated (i.e., both males and females displayed). 

```{r}
active_data<-active_data[active_data$bout_number %in% reciprocated$bout,]
```


We will consider the following effects and perform model selection:

- sex (aka 'subject')
- total time of courtship
- time of day
- day filmed
- group size
- Trial/bout_number (random effect)

```{r modelSelection}
options(na.action = "na.fail")

## full model
model <-lmer(logDuration ~ subject*Group_size + Time_of_Day * Day_filmed + logTotalCourtship + (1|Trial/bout_number), data=active_data)



z <- dredge(model, rank = "AICc")

ztab<-as.data.frame(z)
ztab$model<-gsub("^.*lmer\\(formula = (.*), data.*$","\\1",as.vector(attr(z,"model.calls")))


kable(z, "latex",booktab=TRUE,caption="Model selection table") %>% 
  kable_styling(latex_options=c("HOLD_position", "scale_down")) 

write.csv(ztab,"docs/3_RoyalSocOpenSci/1_R1/Supp1_model_selection_behavior.csv")
```

The top 3 models have very similar AICc scores, and these have subject (1st), subject + group_size (2nd), subject * group_size (3rd). 

```{r BestModel}
model1 <- lmer(logDuration ~ subject + (1|Trial/bout_number), data=active_data)

summary(model1)
anova(model1)

plot(model1)

m1<-emmeans(model1, list("subject"), pbkrtest.limit=3558, type='response')
m1
```


```{r modelTwo}
model2 <- lmer(logDuration ~ subject + Group_size +  (1|Trial/bout_number), data=active_data)

summary(model2)
anova(model2)

m2<-emmeans(model2, list("subject"), pbkrtest.limit=3558, type='response')
m2
```


```{r modelThree}
model3 <- lmer(logDuration ~ subject* Group_size + (1|Trial/bout_number), data=active_data)

summary(model3)
anova(model3)

m3<-emmeans(model3, spec= c('subject','Group_size'), level= 0.95,pbkrtest.limit=3558)
m3
```


```{r modelFour}
model4 <- lmer(logDuration ~ subject + Group_size + Time_of_Day + (1|Trial/bout_number), data=active_data)

summary(model4)
anova(model4)

m4<-emmeans(model4, spec= c('subject','Group_size', 'Time_of_Day'), level= 0.95,pbkrtest.limit=3558)
m4
```


```{r modelFive}
model5 <- lmer(logDuration ~ subject + Time_of_Day + (1|Trial/bout_number), data=active_data)

summary(model5)

m5<-emmeans(model5, spec= c('subject','Time_of_Day'), level= 0.95,pbkrtest.limit=3558)
m5
```

```{r topFive}
top5<-bind_rows(
  mod1=data.frame(m1)[1,],
  mod1=data.frame(m1)[2,],
  mod2=data.frame(m2)[1,],
  mod2=data.frame(m2)[2,],
  mod3=data.frame(m3)[1,],
  mod3=data.frame(m3)[2,],
  mod4=data.frame(m4)[1,],
  mod4=data.frame(m4)[2,],
  mod4=data.frame(m4)[3,],
  mod4=data.frame(m4)[4,],
  mod5=data.frame(m5)[1,],
  mod5=data.frame(m5)[2,],
  mod5=data.frame(m5)[3,],
  mod5=data.frame(m5)[4,],
  .id = "model"
  
  )

write.csv(top5,"active_top5models.csv")
```


For plotting, I'll visualize the durations from females and from males, but also the breakdown across group size. I'll also plot the lines and CIs for males and females across group size, using the function `predicted_probs()` in `predicted_probs.R` (also used in `Chasing_Behaviours.Rmd`).


```{r}
source("R_scripts/predicted_probs.R")
```

```{r}
maldat <- active_data[active_data$subject=="Male",c("subject","Group_size","logDuration", "Trial","bout_number")]
femdat <- active_data[active_data$subject=="Female",c("subject","Group_size","logDuration", "Trial","bout_number")]

jvalues<- seq(from=min(active_data$Group_size), to=max(active_data$Group_size), length.out=100)
femGrpPredict<- predicted_probs(model3, femdat,"logDuration", jvalues)

malGrpPredict<- predicted_probs(model3, maldat,"logDuration", jvalues)

```



```{r activityPlots,fig.keep='last', fig.height=4, fig.width=7.5}
layout(matrix(c(1,2), 1, 2, byrow = TRUE),
   widths=c(2,3))
par(mar=c(4,4,2,1))
# just males and females
vioplot(active_data$logDuration~active_data$subject,
        yaxt='n',
        xlab="",
        ylab="Display duration (s)",
        col=sex_cols,
        axes=FALSE)
axis(1,at=c(1,2),labels=c("Female","Male"),lty=0)

axis(2, at=seq(-6,4,2),
     labels=round(exp(seq(-6,4,2)),3))
legend("topright",
       as.expression(bquote(bold("A"))),
       cex=2,bty='n',xpd=TRUE)


# group size
par(mar=c(4,2,2,2))
plot(active_data$logDuration~active_data$Group_size,
     col=scales::alpha(sex_cols[active_data$subject],0.3),
     cex=1.5,
     xlab="Mean group size",
     ylab="",
     yaxt='n',
     pch=c(1,6)[active_data$subject],
     bty='l')

axis(2, at=seq(-6,4,2),
     labels=round(exp(seq(-6,4,2)),3))

# add lines
polygon(c(rev(femGrpPredict$Variable),femGrpPredict$Variable), 
        c(rev(femGrpPredict$Lower),femGrpPredict$Upper),
     col = alpha("grey",0.5), border = NA)


lines(femGrpPredict$Variable, 
      femGrpPredict$PredictedProbability, lwd = 2,col=sex_cols["Female"])
lines(femGrpPredict$Variable, 
      femGrpPredict$Lower, lwd = 1,col=sex_cols["Female"])
lines(femGrpPredict$Variable, 
      femGrpPredict$Upper, lwd = 1,col=sex_cols["Female"])

# add lines
polygon(c(rev(malGrpPredict$Variable),malGrpPredict$Variable), 
        c(rev(malGrpPredict$Lower),malGrpPredict$Upper),
     col = alpha("grey",0.75), border = NA)


lines(malGrpPredict$Variable, 
      malGrpPredict$PredictedProbability, lwd = 2,col=sex_cols["Male"],lty=2)
lines(malGrpPredict$Variable, 
      malGrpPredict$Lower, lwd = 1,col=sex_cols["Male"],lty=2)
lines(malGrpPredict$Variable, 
      malGrpPredict$Upper, lwd = 1,col=sex_cols["Male"],lty=2)

# add the legend

legend("bottomright",
       c("Female","Male"),
       pch=c(1,6),
       col=sex_cols,
       lty=c(1,2),
       lwd=2,
       cex=1.5,
       bty='n')

legend("topright",
       as.expression(bquote(bold("B"))),
       cex=2,bty='n',xpd=TRUE)
```

## Combine the two major figures


```{r Fig1Combined, out.height='10in',out.width='6in', dpi=300, fig.keep='last'}

# initiation
par(mar=c(4,4,2.5,2.5),
    mfrow=c(2,2))

barplot(recipInit, 
        xlab="",
        ylab="", 
        names.arg=c("Female", "Male"), 
        col=sex_cols[c("Female", "Male")],
        ylim=c(0,200))
legend(y=250,x=-0.4,as.expression(bquote(bold("A"))),cex=2,bty='n',xpd=TRUE)
mtext("Sex initiating courtship", 1,line=2)
mtext("Number of courtship events",2, line=2)

barplot(Comp[,c(2,1)],
        ylab = "",
        xlab = "",
        ylim =c(0,100),
        density=c(0,15,0,15),
        legend=rownames(Comp),
        args.legend=list(
          x="left",
          bty='n'
        ),
        col=sex_cols["Female"]
)
legend(y=125,x=-0.4,as.expression(bquote(bold("B"))),cex=2,bty='n',xpd=TRUE)
mtext("Sex receiving female displays", 1, line=2)
mtext("% of female displays",2,line=2)


# just males and females
vioplot(active_data$logDuration~active_data$subject,
        yaxt='n',
        xlab="",
        ylab="",
        col=sex_cols,
        ylim=c(-1,log(60)),
        axes=FALSE)
axis(1,at=c(1,2),labels=c("Female","Male"),lty=0)

axis(2, at=seq(-6,4,2),
     labels=round(exp(seq(-6,4,2)),1))
mtext("Display duration (s)",2,line=2)
legend(x=0, y=5,
       as.expression(bquote(bold("C"))),
       cex=2,bty='n',xpd=TRUE)


# group size
plot(active_data$logDuration~active_data$Group_size,
     col=scales::alpha(sex_cols[active_data$subject],0.3),
     cex=1.5,
     xlab="Mean group size",
     ylab="",
     yaxt='n',
     ylim=c(-1,log(90)),
     pch=c(1,6)[active_data$subject],
     bty='l')

axis(2, at=seq(-6,4,2),
     labels=round(exp(seq(-6,4,2)),1))
mtext("Display duration (s)",2,line=2)
# add lines
polygon(c(rev(femGrpPredict$Variable),femGrpPredict$Variable), 
        c(rev(femGrpPredict$Lower),femGrpPredict$Upper),
     col = alpha("grey",0.5), border = NA)


lines(femGrpPredict$Variable, 
      femGrpPredict$PredictedProbability, lwd = 2,col=sex_cols["Female"])
lines(femGrpPredict$Variable, 
      femGrpPredict$Lower, lwd = 1,col=sex_cols["Female"])
lines(femGrpPredict$Variable, 
      femGrpPredict$Upper, lwd = 1,col=sex_cols["Female"])

# add lines
polygon(c(rev(malGrpPredict$Variable),malGrpPredict$Variable), 
        c(rev(malGrpPredict$Lower),malGrpPredict$Upper),
     col = alpha("grey",0.75), border = NA)


lines(malGrpPredict$Variable, 
      malGrpPredict$PredictedProbability, lwd = 2,col=sex_cols["Male"],lty=2)
lines(malGrpPredict$Variable, 
      malGrpPredict$Lower, lwd = 1,col=sex_cols["Male"],lty=2)
lines(malGrpPredict$Variable, 
      malGrpPredict$Upper, lwd = 1,col=sex_cols["Male"],lty=2)

# add the legend

legend("topright",
       c("Female","Male"),
       pch=c(1,6),
       col=sex_cols,
       lty=c(1,2),
       lwd=2,
       bty='n')

legend(x=1, y=6,
       as.expression(bquote(bold("D"))),
       cex=2,bty='n',xpd=TRUE)
```


