---
title: "Active courtship analysis"
output: 
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: no
    toc: no
editor_options: 
  chunk_output_type: console
header-includes: |
  \usepackage{lipsum} \usepackage{float} \floatplacement{figure}{H}
---

```{r knitsetup, include=FALSE}
knitr::opts_knit$set(root.dir='../',fig_path="../figs/")
```


```{r chunksetup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.extra='',fig.pos="H",
                      fig.path = "../figs/",
                      dpi = 300,fig.keep='none',dev='png')
```

```{r loadLibs, message=FALSE}
library(stringr)
library(dplyr)
library(scales)
library(tidyverse)
require(lme4)
library(knitr)
library(emmeans)
library(MuMIn)
library(vioplot)
library(kableExtra)
```

```{r}
sex_cols<-c(Female="#7fc97f",Male="#beaed4")
trial_pch<-c(0:7,9:10)
```

```{r}
behav_data<-read.csv("processed_data/combined_behavioural_data.csv")
```

We can summarize the overall behaviours and work out some summary statistics:

```{r}
behav_tab<-table(behav_data$behavior)
```

Of the `r nrow(behav_data)` observed behaviours, `r sum(behav_tab[c("Inactive", "Out of sight")])` are either "Inactive" or "Out of sight". Disregarding those, and our 'active' courtship behaviours make up `r sum(behav_tab[c("Pose", "Wiggle")])/(nrow(behav_data) - sum(behav_tab[c("Inactive", "Out of sight")]))*100`. Chasing and surrounding make up the next largest chunk of behaviours, taking `r sum(behav_tab[c("Chasing", "Surrounding")])/(nrow(behav_data) - sum(behav_tab[c("Inactive", "Out of sight")]))*100`.

```{r}
active_data <- behav_data[behav_data$behavior %in% c("Pose","Wiggle"),]
active_data$logDuration <- log(active_data$Duration)
active_data$logTotalCourtship <- log(active_data$total_time_of_courtship)
```

We don't care about Second female vs female so we should re-code those...

```{r}
active_data$subject[active_data$subject == "Second female"] <- "Female"
```

One question is whether we want to restrict the analysis to behaviours that are intersexual rather than any female-female interactions.... I think we probably do, so let's remove those observations.

```{r}
active_data <- active_data[which(active_data$subject != active_data$modifier_2),]
active_data <- active_data[which(active_data$modifier_2 != "None"),]
```


And we need to ensure that the factors are factors

```{r}
active_data$subject<- factor(active_data$subject, levels=c("Female","Male"))
active_data$Trial <- factor(active_data$Trial, levels=c(
  "Trial 1",
  "Trial 2",
  "Trial 3",
  "Trial 4",
  "Trial 5",
  "Trial 6",
  "Trial 7",
  "Trial 8",
  "Trial 9",
  "Trial 10"
))

active_data$Time_of_Day <- factor(active_data$Time_of_Day, levels=c(
  "AM","Noon"
))
```



## Exploratory analysis

```{r}
table(active_data$behavior, active_data$subject)
```




```{r}
par(mfrow=c(2,2))
hist(active_data$Duration)
hist(active_data$logDuration)
hist(active_data$total_time_of_courtship)
hist(active_data$logTotalCourtship)
```




What are the things that we want to investigate? 

- sex (aka 'subject')
- total time of courtship
- time of day
- day filmed
- group size
- Trial/bout_number (random effect)

```{r}
summary(active_data[,c("logDuration","subject","Time_of_Day","Day_filmed","total_time_of_courtship","Group_size","Trial","bout_number")])
```


are any of these highly correlated?


```{r}
par(mfrow=c(1,5))
plot(as.factor(active_data$Trial), as.factor(active_data$subject))
plot(as.factor(active_data$Trial), active_data$Group_size)
plot(as.factor(active_data$Trial), active_data$Day_filmed)
plot(as.factor(active_data$Trial), active_data$total_time_of_courtship)
plot(as.factor(active_data$Trial), as.factor(active_data$Time_of_Day))
```

```{r}
par(mfrow=c(1,5))
plot(active_data$bout_number, as.factor(active_data$subject))
plot(active_data$bout_number, active_data$Group_size)
plot(active_data$bout_number, active_data$Day_filmed)
plot(active_data$bout_number, active_data$total_time_of_courtship)
plot(active_data$bout_number, as.factor(active_data$Time_of_Day))
```

```{r}
par(mfrow=c(1,4))
plot(active_data$Group_size, as.factor(active_data$subject))
plot(active_data$Group_size, active_data$Day_filmed)
plot(active_data$Group_size, active_data$total_time_of_courtship)
plot(active_data$Group_size, as.factor(active_data$Time_of_Day))
```

```{r}
par(mfrow=c(1,3))
plot(active_data$Day_filmed, as.factor(active_data$subject))
plot(active_data$Day_filmed, active_data$total_time_of_courtship)
plot(active_data$Day_filmed, as.factor(active_data$Time_of_Day))
```

```{r}
par(mfrow=c(1,2))
plot(active_data$total_time_of_courtship, as.factor(active_data$subject))
plot(active_data$total_time_of_courtship, as.factor(active_data$Time_of_Day))
```


```{r}
par(mfrow=c(1,1))
plot(as.factor(active_data$subject), as.factor(active_data$Time_of_Day))
```

Amazingly there don't seem to be any major relationships between these predictors. 


## Model selection

- sex (aka 'subject')
- total time of courtship
- time of day
- day filmed
- group size
- Trial/bout_number (random effect)

```{r modelSelection}
options(na.action = "na.fail")

## full model
model <-lmer(logDuration ~ subject*Group_size + Time_of_Day * Day_filmed + logTotalCourtship + (1|Trial/bout_number), data=active_data)



z <- dredge(model, rank = "AICc")

kable(z, "latex",booktab=TRUE,caption="Model selection table")
```

The top 3 models have very similar AICc scores, and these have subject (1st), subject + group_size (2nd), subject * group_size (3rd). 

```{r BestModel}
model1 <- lmer(logDuration ~ subject + (1|Trial/bout_number), data=active_data)

summary(model1)
anova(model1)

plot(model1)

emmeans(model1, list("subject"), pbkrtest.limit=3558)
```


```{r}
model2 <- lmer(logDuration ~ subject + Group_size +  (1|Trial/bout_number), data=active_data)

summary(model2)



```


```{r}
model3 <- lmer(logDuration ~ subject* Group_size + (1|Trial/bout_number), data=active_data)


```


```{r}
layout(matrix(c(1,2), 1, 2, byrow = TRUE),
   widths=c(2,3))
par(mar=c(4,4,2,1))
# just males and females
vioplot(active_data$logDuration~active_data$subject,
        yaxt='n',
        xlab="",
        ylab="Display duration (s)",
        col=sex_cols)
axis(1,at=c(1,2),labels=c("Female","Male"))

axis(2, at=seq(-6,4,2),
     labels=round(exp(seq(-6,4,2)),3))


# group size
par(mar=c(4,2,2,2))
plot(active_data$logDuration~active_data$Group_size,
     col=sex_cols[active_data$subject],
     cex=1.5,
     xlab="Mean group size",
     ylab="",
     yaxt='n')

axis(2, at=seq(-6,4,2),
     labels=round(exp(seq(-6,4,2)),3))
```





