---
title: "Combining data sheets for analysis"
output: 
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: no
    toc: no
editor_options: 
  chunk_output_type: console
header-includes: |
  \usepackage{lipsum} \usepackage{float} \floatplacement{figure}{H}
---

```{r knitsetup, include=FALSE}
knitr::opts_knit$set(root.dir='../',fig_path="../figs/")
```


```{r chunksetup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.extra='',fig.pos="H",
                      fig.path = "../figs/",
                      dpi = 300,fig.keep='none',dev='png')
```

```{r loadLibs, message=FALSE}
library(stringr)
library(dplyr)
library(scales)
library(tidyverse)
require(lme4)
library(knitr)
library(VGAM)
```

```{r}
time_cols<-c(AM="#fc8d62",Noon="#e78ac3")
trial_pch<-c(0:7,9:10)
```
```{r}
logit_transform<-function(x){
  xl<-log(x/(1-x))
  return(xl)
}
```

The code in this document is modified from Fleur's scripts `Readin_data.R` and `Combining_data.R`.  The code to parse the chasing data is taken and modified from Fleur's script `Chasing_data.R` -- these R scripts have been moved to an archive because this document contains the code used for the final analysis. All of the courtship behaviours are in `BORIS_data/` and the chase datasheets are in `Chase_datasheets/`.
 
## Parsing the data

### Courtship data 

```{r readData}
courtfiles<-list.files(path="BORIS_data/",pattern="csv",full.names=TRUE)


courtdat<- lapply(courtfiles, function(file){
  #scan through data to line ending 
  temp<-scan(file, "character", sep='\n', quiet = TRUE) 
   #find line that starts with Time
  nlines<- grep("Time,", temp)                     
  # skip everything above line that starts with Time
  dat<-read.csv(file, skip=nlines-1) 
  # Mod 3 has both numbers and words change to character
  dat$Modifier.3<-as.character(dat$Modifier.3) 
  dat$file.name<-file         
  return(dat)
})   
courtdat<-bind_rows(courtdat)

```

```{r extractTimesFXN}
# This function will extract the times for a behaviour's start and end points
extract_times <- function(dat1){
  
  nn <- nrow(dat1)
  
  starting_points <- c()
  end_points <- c()
  
  #get where observations start
  for(i in 1:nn){
    d <- dat1[i,]
    
    behavior <- d$Behavior
    
    if(behavior == "Start"){
      #print("start")
      starting_points <- c(starting_points, i)
    }
    
    if(behavior == "Stop"){
      #print("end")
      end_points <- c(end_points, i)
    }
    
    
  }
  return(list(starting_points,end_points))
}
  
```

```{r extractTimes}
times<-extract_times(courtdat) 
starting_points<-times[[1]]
end_points<-times[[2]]
observation_chunks <- list()

 for(j in 1:length(starting_points) ){
    start <- starting_points[j]
    end <- end_points[j]
    observation_chunks[[j]] <- courtdat[start:end, ]
    
  }
```

```{r getUsefulInfo}
courtdat_list <- lapply(seq_along(observation_chunks), function(chunk){
  
  x <- observation_chunks[[chunk]]
  x$Modifier.1[is.na(x$Modifier.1)]<-""
  x$Modifier.2[is.na(x$Modifier.2)]<-""
  x$Modifier.3[is.na(x$Modifier.3)]<-""
  n3 <- nrow(x)
  
  length_of_chunk <- x$Time[n3] - x$Time[1] 
  times <-data.frame()
  
  for(k in 1:n3){
    x_row <- x[k,]
    
    if(x_row$Behavior== "Start" | x_row$Behavior== "Stop" ){
      next()
    }else{
      
      selected_behavior <- x_row$Behavior
      selected_subject <- x_row$Subject
      selected_m1 <- x_row$Modifier.1
      selected_m2 <- x_row$Modifier.2 
      selected_m3 <- x_row$Modifier.3
      selected_status <- x_row$Status
      
      if(selected_status=="START"){
        end_row <- x[which   (x$Behavior==selected_behavior &
                             x$Subject== selected_subject & 
                             x$Modifier.1 == selected_m1 &
                             x$Modifier.2 == selected_m2 &
                             x$Modifier.3 == selected_m3 &
                             x$Status == "STOP" &
                               x$Time>=x_row$Time),]
       
       
        time_elapsed <- end_row$Time[which.min(end_row$Time-x_row$Time)] - x_row$Time
        #}
        if(nrow(end_row)==0){
          print(paste(x_row$Media.file.path,x_row$Time,time_elapsed ))
        }
    
        return_data_per_row <- data.frame("media_file_path"= x_row$Media.file.path,
                                          "time_in_video"= x_row$Time,
                                          "Duration"= time_elapsed,
                                          "behavior"= selected_behavior,
                                          "subject"=selected_subject,
                                          "modifier_1" = selected_m1, 
                                          "modifier_2"= selected_m2,
                                          "modifier_3"= selected_m3)
        
        
        times <-rbind(times, return_data_per_row )
        
      }
      
      if(selected_status == "STOP"){
        next()
      }
      
      if(selected_status == "POINT"){
       
        return_data_per_row <- data.frame("media_file_path"= x_row$Media.file.path,
                                          "time_in_video"= x_row$Time,
                                          "Duration"= 0.00,
                                          "behavior"= selected_behavior,
                                          "subject"=selected_subject,
                                          "modifier_1" = selected_m1, 
                                          "modifier_2"= selected_m2,
                                          "modifier_3"= selected_m3)
        times <-rbind(times, return_data_per_row )
      }
    }
  }
  times$bout_number <- chunk
  times$total_time_of_courtship <- length_of_chunk
  
  return(times)  
})

court_data <- do.call(rbind, courtdat_list)
court_data$Trial=substr(court_data$media_file_path, 27, 33)
#Adding Am or Noon filming times
court_data$Time_of_Day=str_sub(court_data$media_file_path, -10)
court_data$Time_of_Day=str_extract(court_data$Time_of_Day,"(\\w+)")
court_data$Time_of_Day <-  ifelse(court_data$Time_of_Day == "oon", "Noon",
                                  court_data$Time_of_Day)
#Adding Date filmed 
court_data$Date_filmed=substr(court_data$media_file_path, 35, 42)
court_data$Date_filmed=str_extract(court_data$Date_filmed,"(\\w+)")

# Adding which day filmed (1-7)
court_data$Day_filmed <- ifelse(court_data$Trial == "Trial 1" & court_data$Date_filmed == 8, "7",
                         ifelse(court_data$Trial == "Trial 2" & court_data$Date_filmed == 3, "1",
                         ifelse( court_data$Trial == "Trial 2" & court_data$Date_filmed == 9, "7",
                         ifelse( court_data$Trial == "Trial 3" & court_data$Date_filmed == 3, "1",
                         ifelse( court_data$Trial == "Trial 3" & court_data$Date_filmed == 4, "2",
                         ifelse( court_data$Trial == "Trial 3" & court_data$Date_filmed == 5, "3",
                         ifelse( court_data$Trial == "Trial 3" & court_data$Date_filmed == 8, "6",
                         ifelse( court_data$Trial == "Trial 4" & court_data$Date_filmed == 4, "1",
                         ifelse( court_data$Trial == "Trial 4" & court_data$Date_filmed == 5, "2",
                         ifelse( court_data$Trial == "Trial 4" & court_data$Date_filmed == 6, "3",
                         ifelse( court_data$Trial == "Trial 4" & court_data$Date_filmed == 7, "4",
                         ifelse( court_data$Trial == "Trial 4" & court_data$Date_filmed == 10, "7",
                         ifelse( court_data$Trial == "Trial 5" & court_data$Date_filmed == 6, "2",
                         ifelse( court_data$Trial == "Trial 5" & court_data$Date_filmed == 9, "5",
                         ifelse( court_data$Trial == "Trial 5" & court_data$Date_filmed == 10, "6",                    
                         ifelse( court_data$Trial == "Trial 5" & court_data$Date_filmed == 11, "7",
                         ifelse( court_data$Trial == "Trial 6" & court_data$Date_filmed == 7, "3",
                         ifelse( court_data$Trial == "Trial 6" & court_data$Date_filmed == 8, "4",
                         ifelse( court_data$Trial == "Trial 6" & court_data$Date_filmed == 10, "6",
                         ifelse( court_data$Trial == "Trial 7" & court_data$Date_filmed == 9, "2",
                         ifelse( court_data$Trial == "Trial 7" & court_data$Date_filmed == 10, "3",
                         ifelse( court_data$Trial == "Trial 7" & court_data$Date_filmed == 13, "6",
                         ifelse( court_data$Trial == "Trial 8" & court_data$Date_filmed == 8, "1",
                         ifelse( court_data$Trial == "Trial 8" & court_data$Date_filmed == 10, "3",
                         ifelse( court_data$Trial == "Trial 8" & court_data$Date_filmed == 11, "4",
                         ifelse( court_data$Trial == "Trial 8" & court_data$Date_filmed == 13, "6",
                         ifelse( court_data$Trial == "Trial 8" & court_data$Date_filmed == 14, "7",
                         ifelse( court_data$Trial == "Trial 9" & court_data$Date_filmed == 10, "2",
                         ifelse( court_data$Trial == "Trial 9" & court_data$Date_filmed == 11, "3",
                         ifelse( court_data$Trial == "Trial 9" & court_data$Date_filmed == 13, "5",
                         ifelse( court_data$Trial == "Trial 10" & court_data$Date_filmed == 12, "3",
                         NA  )))))))))))))))))))))))))))))))

#Trial 10 
court_data$Trial <- ifelse(court_data$Trial == "Trial 1" & court_data$Date_filmed == 12, "Trial 10",
                           court_data$Trial)
court_data$Day_filmed <- ifelse( court_data$Trial == "Trial 10" & court_data$Date_filmed == 12, "3",
                                 court_data$Day_filmed)

  
```
```{r saveCourtData}
write.csv(court_data,"processed_data/courtship_data.csv", row.names=FALSE)
```



### Chasing data

```{r getData}
chase_data<-dplyr::bind_rows(
  lapply(
    list.files(
      pattern="Trial", 
      path="Chase_datasheets",
      full.names = TRUE)
    , read.csv
    )
  )
colnames(chase_data) <- c("media_file_path", 
                          "time_in_video", 
                          "Duration",  
                          "behavior" , 
                          "subject" ,
                          "modifier_1" ,
                          "modifier_2",
                          "modifier_3" ,
                          "Group_size", 
                          "Female", 
                          "Male" ,
                          "Chasing_end" , 
                          "Chasing_present" , 
                          "bout_number_wrong" ,
                          "total_time_of_courtship")

```

```{r cleanData, warning=FALSE}
# remove extra columns
chase_data<-chase_data[,1:which(colnames(chase_data)=="total_time_of_courtship")]

# remove bouts with no active courtship
# aka female-female interactions where group size was NA
chase_data<- subset(chase_data,!is.na(Group_size)) 
chase_data <- chase_data %>%
  mutate_at(vars(Group_size,Female,Male), funs(round(., 1))) 

# remove column with incorrect bout numbers
chase_data <- chase_data[,-which(colnames(chase_data)=="bout_number_wrong")]
```

```{r metaData}
#Adding in Trial number
chase_data$Trial=substr(chase_data$media_file_path, 27, 33)
#Adding Am or Noon filming times
chase_data$Time_of_Day=str_sub(chase_data$media_file_path, -10)
chase_data$Time_of_Day=str_extract(chase_data$Time_of_Day,"(\\w+)")
chase_data$Time_of_Day <-  ifelse(chase_data$Time_of_Day == "oon", "Noon",
                                  chase_data$Time_of_Day)
#Adding Date filmed 
chase_data$Date_filmed=substr(chase_data$media_file_path, 35, 42)
chase_data$Date_filmed=str_extract(chase_data$Date_filmed,"(\\w+)")

# Adding which day filmed (1-7)
chase_data$Day_filmed <- ifelse(chase_data$Trial == "Trial 1" & chase_data$Date_filmed == 8, "7",
                         ifelse(chase_data$Trial == "Trial 2" & chase_data$Date_filmed == 3, "1",
                         ifelse( chase_data$Trial == "Trial 2" & chase_data$Date_filmed == 9, "7",
                         ifelse( chase_data$Trial == "Trial 3" & chase_data$Date_filmed == 3, "1",
                         ifelse( chase_data$Trial == "Trial 3" & chase_data$Date_filmed == 4, "2",
                         ifelse( chase_data$Trial == "Trial 3" & chase_data$Date_filmed == 5, "3",
                         ifelse( chase_data$Trial == "Trial 3" & chase_data$Date_filmed == 8, "6",
                         ifelse( chase_data$Trial == "Trial 4" & chase_data$Date_filmed == 4, "1",
                         ifelse( chase_data$Trial == "Trial 4" & chase_data$Date_filmed == 6, "3",
                         ifelse( chase_data$Trial == "Trial 4" & chase_data$Date_filmed == 7, "4",
                         ifelse( chase_data$Trial == "Trial 4" & chase_data$Date_filmed == 10, "7",
                         ifelse( chase_data$Trial == "Trial 5" & chase_data$Date_filmed == 6, "2",
                         ifelse( chase_data$Trial == "Trial 5" & chase_data$Date_filmed == 9, "5",
                         ifelse( chase_data$Trial == "Trial 5" & chase_data$Date_filmed == 10, "6",                    
                         ifelse( chase_data$Trial == "Trial 5" & chase_data$Date_filmed == 11, "7",
                         ifelse( chase_data$Trial == "Trial 6" & chase_data$Date_filmed == 7, "3",
                         ifelse( chase_data$Trial == "Trial 6" & chase_data$Date_filmed == 8, "4",
                         ifelse( chase_data$Trial == "Trial 6" & chase_data$Date_filmed == 10, "6",
                         ifelse( chase_data$Trial == "Trial 7" & chase_data$Date_filmed == 10, "3",
                         ifelse( chase_data$Trial == "Trial 7" & chase_data$Date_filmed == 13, "6",
                         ifelse( chase_data$Trial == "Trial 8" & chase_data$Date_filmed == 10, "3",
                         ifelse( chase_data$Trial == "Trial 8" & chase_data$Date_filmed == 11, "4",
                         ifelse( chase_data$Trial == "Trial 8" & chase_data$Date_filmed == 13, "6",
                         ifelse( chase_data$Trial == "Trial 8" & chase_data$Date_filmed == 14, "7",
                         ifelse( chase_data$Trial == "Trial 9" & chase_data$Date_filmed == 10, "2",
                         ifelse( chase_data$Trial == "Trial 9" & chase_data$Date_filmed == 11, "3",
                         ifelse( chase_data$Trial == "Trial 9" & chase_data$Date_filmed == 13, "5",
                         ifelse( chase_data$Trial == "Trial 10" & chase_data$Date_filmed == 12, "3",
                         NA  ))))))))))))))))))))))))))))

#Trial 10 - because R though it was trail 1
chase_data$Trial <- ifelse(chase_data$Trial == "Trial 1" & chase_data$Date_filmed == 12, "Trial 10",
                           chase_data$Trial)
chase_data$Day_filmed <- ifelse( chase_data$Trial == "Trial 10" & chase_data$Date_filmed == 12, "3",
                                 chase_data$Day_filmed)



```



```{r femaleSize}
chase_data <- chase_data %>% mutate(Max_length =
                     case_when(Trial =="Trial 1" ~ "115.2", 
                               Trial =="Trial 2" ~ "116.6",
                               Trial =="Trial 3" ~ "121.4",
                               Trial =="Trial 4" ~ "123.6",
                               Trial =="Trial 5" ~ "113.2",
                               Trial =="Trial 6" ~ "119.4",
                               Trial =="Trial 7" ~ "122.2",
                               Trial =="Trial 8" ~ "113.7",
                               Trial =="Trial 9" ~ "116.4",
                               Trial =="Trial 10"~ "113.5",                   
                               ))

chase_data <- chase_data %>% mutate(Min_length =
                               case_when(Trial =="Trial 1" ~ "87.8", 
                                         Trial =="Trial 2" ~ "89.8",
                                         Trial =="Trial 3" ~ "90.4",
                                         Trial =="Trial 4" ~ "86.8",
                                         Trial =="Trial 5" ~ "94.2",
                                         Trial =="Trial 6" ~ "83.2",
                                         Trial =="Trial 7" ~ "83.9",
                                         Trial =="Trial 8" ~ "88",
                                         Trial =="Trial 9" ~ "94.8",
                                         Trial =="Trial 10"~ "94.4",                   
                               ))

chase_data[ , c("Max_length","Min_length")]<- apply(
  chase_data[ , c("Max_length","Min_length")], 
  2,
  function(x) as.numeric(as.character(x))
)
chase_data$Difference<- (chase_data$Max_length - chase_data$Min_length)

```
```{r chasingNum}
chase_data$Chasing_numeric<-ifelse(chase_data$Chasing_present=="Yes",1, 0)
chase_data$Chasing_end_numeric<-ifelse(chase_data$Chasing_end=="Yes",1, 0)
```
```{r saveChaseData}
write.csv(chase_data, "processed_data/chase_data.csv", row.names = FALSE)
```



### Combining courtship and chasing data

```{r}
court_data<-read.csv("processed_data/courtship_data.csv")
chase_data<-read.csv("processed_data/chase_data.csv")
```

```{r createUniqueID}
court_data$combo<-paste(gsub("\\s+","",court_data$media_file_path), 
                        court_data$time_in_video, 
                        round(court_data$Duration,3), 
                        round(court_data$total_time_of_courtship,3), sep="_")
court_data<-unique(court_data)

chase_data$combo<-paste(gsub("\\s+","",chase_data$media_file_path), 
                      chase_data$time_in_video, 
                      round(chase_data$Duration,3) ,
                      round(chase_data$total_time_of_courtship,3), sep="_")
chase_data<-unique(chase_data)
chase_data$Trial<-factor(chase_data$Trial,levels=c(
  "Trial 1",
  "Trial 2",
  "Trial 3",
  "Trial 4",
  "Trial 5",
  "Trial 6",
  "Trial 7",
  "Trial 8",
  "Trial 9",
  "Trial 10"
))

behav_data <- merge(court_data, chase_data[,c("combo","Group_size", "Female", "Male","Chasing_present","Chasing_numeric","Chasing_end","Chasing_end_numeric")], 
by="combo")
# remove any unnecessarily duplicated ones
behav_data<-unique(behav_data)

# re-order each bout based on time points
behav_data<-do.call(rbind,by(behav_data,behav_data$bout_number,function(bout_dat){
  sort_dat<-bout_dat[order(bout_dat$time_in_video),]
  return(sort_dat)
}))
```

So, there are some interesting issues. There are fewer chasing observations than courtship observations, but this is expected -- chasing only occurred in intersexual contexts, and the courtship data contains some same-sex interactions. The courtship dataset also contains a few observations without active courtship from all participants, which were excluded from analyses.

But there are also some courtship bouts that are missing from the merged dataframe  -- let's take a look at these.

```{r checkingMissing}
missing<-court_data[!court_data$combo %in% behav_data$combo,]

intersex<-unlist(as.list(by(missing,missing$bout_number,function(bout){
  
  males<- bout$subject %in% "Male"
  
  if(length(males[males==TRUE]) == 0){
    return(FALSE)
  } else{
    return(TRUE)
  }
  
})))

# these are the bouts that should be included
#intersex[intersex==TRUE]
missing<-missing[missing$bout_number %in% names(intersex[intersex==TRUE]),]

active_behavs<-c("Wiggle","Pose")

active<-unlist(as.list(by(missing, missing$bout_number, function(bout, active_behavs){
  
  activeTF<-bout$behavior %in% active_behavs
  if(length(activeTF[activeTF==TRUE])==0){
    return(FALSE)
  }else {
    return(TRUE)
  }
},active_behavs=active_behavs)))
active_bouts<-active[active==TRUE]

missing<-missing[missing$bout_number %in% names(active_bouts),]
table(missing$media_file_path, missing$bout_number)
```

No courtship bouts remain that should be in the chase data but are not -- all courtship events involving both males and females and active behaviours are present.


```{r, eval=FALSE, include=FALSE}
prob_file<-unique(chase_data$media_file_path[gsub("\\s+","",chase_data$media_file_path) %in% gsub("\\s+","",unique(missing$media_file_path))])

# number of observations in that trial
n_court<-nrow(court_data[gsub("\\s+","",court_data$media_file_path) %in% gsub("\\s+","",prob_file),])
n_chase<-nrow(chase_data[gsub("\\s+","",chase_data$media_file_path) %in% gsub("\\s+","",prob_file),])

prob_court<-court_data[gsub("\\s+","",court_data$media_file_path) %in% gsub("\\s+","",prob_file),]
prob_chase<-chase_data[gsub("\\s+","",chase_data$media_file_path) %in% gsub("\\s+","",prob_file),]

missing_times<-prob_court$time_in_video[!prob_court$time_in_video %in% prob_chase$time_in_video]
```

```{r, eval=FALSE, include=FALSE}
hist(prob_chase$time_in_video[prob_chase$time_in_video >= min(missing_times) & prob_chase$time_in_video <= max(missing_times) ],
     col="grey",
     border = "grey",
     ylim=c(0,4),
     xlab="Time in the video",
     main="",
     ylab="Number of observations",
     breaks=seq(floor(min(missing_times))-5,ceiling(max(missing_times))+5,5))
hist(missing_times,add=TRUE, 
     col=alpha("cornflowerblue",0.5),
     border=alpha("cornflowerblue",0.5),
      breaks=seq(floor(min(missing_times))-5,ceiling(max(missing_times))+5,5))
legend("topleft",
       c("Timepoints in the chasing dataset",
         "Timepoints that are in the courting dataset 
and missing from chasing"),
       col=c("grey",alpha("cornflowerblue",0.5)),
       pch=15,
       bty='n',
       ncol=1)
```
```{r}
write.csv(behav_data, "processed_data/combined_behavioural_data.csv", row.names = FALSE)
```

